---
title: "CDS Exam - Analysing Højskolesangbogen"
author: "Jakob Grøhn Damgaard"
date: "12/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages('pacman')
library(pacman)
pacman::p_load(tidyverse, dplyr, gganimate, gifski)
# 'gifski' needed for rendering of animations 
```

## Analysing semantic and linguistic patterns in the Danish Højskolesangbog
This markdown script contains analysis of semantic and linguistic patterns in songs in the 19th version of the Danish Højskolesangbogen. These patterns are then represented using graphical animations and visualisations.
The loaded data is obtained through automated web scraping of hojskolesangbogen.dk performed in Python 3.8.5

**1. Loading and  preparing data**
```{r Loading and  preparing data, results='hide', warning = FALSE, message = FALSE}
# Loading song data
song_data <- read_csv('Data/clean_song_data_with_word_counts.csv')

# Turning 'no data' cells in columns containing years into NAs
song_data$year_written <- as.numeric(song_data$year_written)
song_data$year_composed <- as.numeric(song_data$year_written)

# Adding empty column with time interval
song_data$time_written_interval <- NA

# Moving this column to front of data frame
song_data <- song_data %>%
  select(time_written_interval, everything())


# Loading vocabulary
vocab <- read_csv('Data/song_vocabulary.csv')

# Ordering vocabulary according to designated word numbers
vocab <- vocab[order(vocab$Word),]  

# Creating list of words in vocabulary
vocab_words <- as.vector(vocab$Vocab)

# Changing column names in song_data to real words
colnames(song_data)[9:13326] <- vocab_words


# Create list of most common religious words
rel_words <- c('jesu', 'jesubarnet', 'helligånden', 'hellig', 'gud', 'himmel', 'himlen', 'krist', 'himmelsk', 'fader','herre', 'herren', 'velsigne', 'ånd', 'åndelig', 'rige','himmerrige', 'halleluja', 'guds', 'kristendom','kristelighed','frelserens','velsignelsen', 'velsignet', 'velsigner', 'paradis','jul')
```



**2. Matching religious words with vocabulary**
```{r Find religious words in vocabulary}
# Create empty data frame
rel_words_present <- c()

# Loop through the vocabulary to find the numbers matching the religious words
for(i in colnames(song_data)){
  if (i %in% rel_words){
    rel_words_present = c(rel_words_present, i)
  }
}

# Create empty data frame
rel_words_sums = tibble()


# Find the frequency of the religious words
for(i in as.vector(rel_words_present)){
  rel_words_sums = rbind(rel_words_sums, sum(song_data[,i]))
}

# Creating concatenated data frame
rel_words_overall = cbind(rel_words_sums,rel_words_present)
# Renaming column containing counts
rel_words_overall <- rel_words_overall %>% rename( count = X39)

# Number of words in the corpus
sum(song_data[, 9:13326])
``` 
Songs are categorised into following time intervals for temporal analyses:\
 1. < 1700\
 2. 17-1800\
 3. 1800-1850\
 4. 1850-1900\
 5. 1900-1950\
 5. 2000 >\
\
**3. Temporally categorising songs**
```{r  Temporally categorising songs}
# Grouping songs according to year writting
for(i in 1:nrow(song_data)){
  if (is.na(song_data$year_written[i])){
    song_data$time_written_interval[i] = 'Uknown date'
  } else if (song_data$year_written[i] < 1800){
    song_data$time_written_interval[i] = '< 1800'
  } else if (song_data$year_written[i] >= 1800 & song_data$year_written[i]< 1850){
    song_data$time_written_interval[i] = '1800-1850'
  } else if (song_data$year_written[i] >= 1850 & song_data$year_written[i]< 1900){
    song_data$time_written_interval[i] = '1850-1900'
  } else if (song_data$year_written[i] >= 1900 & song_data$year_written[i]< 1950){
    song_data$time_written_interval[i] = '1900-1950'
  } else if (song_data$year_written[i] >= 1950 & song_data$year_written[i]< 2000){
    song_data$time_written_interval[i] = '1950-2000'
  } else {
    song_data$time_written_interval[i] = '2000 >'
  }
}    

# Define non-numeric columns
drop_cols = c('X1', 'year_written', 'songwriter', 'composer', 'year_composed', 'lyrics', 'title')

# Find number of occurrences of religious words for each time interval
time_grouped_data <- song_data %>%
  select(-one_of(drop_cols)) %>% 
  group_by(time_written_interval) %>% 
  summarise_each(funs(sum)) 
  
# Adding column containing total words in all lyrics in that time interval
time_grouped_data$total_words <-NA
  
# Moving this column to front of data frame
time_grouped_data <- time_grouped_data %>%
  select(total_words, everything())

# Fill column with value outputted by summing all columns 
for (i in 1:nrow(time_grouped_data)){
  time_grouped_data$total_words[i] = sum(time_grouped_data[i,3:13320])
}

# Keep only columns with religious words, total words and category
time_grouped_data <- time_grouped_data %>% select(one_of(rel_words_present) | total_words | time_written_interval)

# Adding column containing total words in all lyrics in that time interval
time_grouped_data$total_rel_words <-NA
  
# Fill column with value outputted by summing all columns 
for (i in 1:nrow(time_grouped_data)){
  time_grouped_data$total_rel_words[i] = sum(time_grouped_data[i,1:21])
}

# Add column contaning frequency of religious words
time_grouped_data$rel_word_freq = time_grouped_data$total_rel_words / time_grouped_data$total_words * 100

# Create empty column for number of songs in each time span
time_grouped_data$total_songs <- as.vector(table(song_data$time_written_interval))

# Remove row with song with unknown data
time_grouped_data <- time_grouped_data[-c(7), ]
``` 


**4. Plots and visualisation of interval aggregated data**
```{r Plots}
# Add column for use as transition time variable in gganimate plot
time_grouped_data$cout <- c(1:6)

# Create line plot of development in frequency of religious words
interval_plot <- ggplot(time_grouped_data, aes(x = as.factor(time_written_interval), y = rel_word_freq, group = 1)) +
  geom_line(color = '#586CB0', size = 1)+
  geom_point(color = '#586CB0', size = 2) +
  labs(x = "Year of writing", # Defining all labels
       y = "Frequency of religious words (% of all words)", 
       title = "Change in religious intensity")+
  theme_bw() +  # Setting theme
  transition_reveal(cout)

# Define plot size and save GIF to working directory
animate(interval_plot, height = 500, width = 750)
anim_save("Visualisations/interval_plot.gif")
```

**5. Get religious intensity for each song**
```{r Religious intensity for each song*}
indivudal_song_data <- song_data[order(song_data$year_written),]  

individual_drop_cols = c('X1', 'songwriter', 'composer', 'year_composed', 'lyrics', 'title')

indivudal_song_data <- indivudal_song_data %>%
  select(-one_of(individual_drop_cols))

# Adding column containing total words in all lyrics in that time interval
indivudal_song_data$total_words <-NA
  
# Moving this column to front of data frame
indivudal_song_data <- indivudal_song_data %>%
  select(total_words, everything())

# Fill column with value outputted by summing all columns 
for (i in 1:nrow(indivudal_song_data)){
  indivudal_song_data$total_words[i] = sum(indivudal_song_data[i,4:13320])
}

# Keep only columns with religious words, total words and category
indivudal_song_data <- indivudal_song_data %>% select(one_of(rel_words_present) | total_words | year_written)

# Adding column containing total words in all lyrics in that time interval
indivudal_song_data$total_rel_words <-NA
  
# Fill column with value outputted by summing all columns 
for (i in 1:nrow(indivudal_song_data)){
  indivudal_song_data$total_rel_words[i] = sum(indivudal_song_data[i,1:21])
}

# Add column contaning frequency of religious words
indivudal_song_data$rel_word_freq = indivudal_song_data$total_rel_words / indivudal_song_data$total_words * 100

# Remove songs with uknown data
indivudal_song_data <- indivudal_song_data %>% filter(is.na(year_written) == FALSE)

# Remove songs with no appertaining texy
indivudal_song_data <- indivudal_song_data %>% filter(total_words != 0)

# Removing outlier songs written before 1500
indivudal_song_data <- indivudal_song_data %>% filter(year_written > 1500)
```

**6. Visualising individual songs**
```{r Plots - 2}
# Create line plot of development in frequency of religious words for each song
ggplot(indivudal_song_data, aes(x = year_written, y = rel_word_freq, group = 1)) +
  geom_line(color = '#586CB0', size = 1)+
  geom_point(color = '#586CB0', size = 2) +
  labs(x = "Year of writing", # Defining all labels
       y = "Frequency of religious words (% of all words)", 
       title = "Change in religious intensity")+
  theme_bw() # Setting theme
  

# Finding most religious song
most_religious_song <- song_data %>% filter(year_written == 1845 & krist == 4)
```

**7. Get religious intensity for an aggregation of every 15th song**
```{r Get religious intensity for an aggregation of every 15th song}
# Creating a copy of data and moving rel_word_freq
song_data_15_collapsed <- indivudal_song_data %>% select(-rel_word_freq)

# Turning year_written values numeric
song_data_15_collapsed$year_written <- as.numeric(song_data_15_collapsed$year_written)


# Aggregating values for every five song 
# Credit: https://stackoverflow.com/questions/30359427/calculate-the-mean-of-every-13-rows-in-data-frame
n <-  15
song_data_15_collapsed <-  aggregate(song_data_15_collapsed,list(rep(1:(nrow(song_data_15_collapsed)%/%n+1),each=n,len=nrow(song_data_15_collapsed))),mean)[-1]

# Creating new column with religious word frequency
song_data_15_collapsed$rel_word_freq = song_data_15_collapsed$total_rel_words / song_data_15_collapsed$total_words * 100
```


**8. Visualising religious intensity for an aggregation of every 15th song**
```{r Plots - 3}
# Creating animated line plot of development in frequency of religious words
ggplot(song_data_15_collapsed, aes(x = year_written, y = rel_word_freq, group = 1)) +
  geom_line(color = '#586CB0', size = 1)+
  geom_point(color = '#586CB0', size = 2) +
  labs(x = "Year of writing", # Defining all labels
       y = "Frequency of religious words (% of all words)", 
       title = "Change in religious intensity") +
  theme_bw() + # Setting theme
  transition_reveal(year_written)

# Saving animation
anim_save("Visualisations/aggregated_songs_15_religious_intensity.gif")

# Plot with smooth line
aggregated_15 <- ggplot(song_data_15_collapsed, aes(x = year_written, y = rel_word_freq, group = 1),
                        width = 750, height = 500) +
  geom_smooth(color = '#586CB0', size = 1, span = 0.5, alpha = 0.15)+
  geom_point(color = '#586CB0', size = 2, alpha = 0.5) +
  labs(x = "Year of writing", # Defining all labels
       y = "Frequency of religious words (% of all words)", 
       title = "Change in Religious intensity")+
  theme_bw() # Setting theme
aggregated_15

# Saving final plot as png
ggsave("Visualisations/aggregated_15.png", width = 6, height = 4)
```

